---
title: "Multi‑Species Indicator (MSI) Workflow"
author: "Dylan"
output: HTML
---

# Setup

```{r}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)

library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)
library(lubridate)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

# Trend start year
trend_start_year = 1980
trend_end_year = 2024

```

# Ants - sparta and occti

## Load OCCTI outputs

```{r}

# Forgot to implement 50 record filter.
all_sp_ants = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
count(species) %>%
filter(n >= 50) %>%
pull(species)

occti_ants_all_sp_outputs <- load_occti_outputs(
  results_dir = "past_jasmin_datalabs_runs/_rslurm_explore_occ_run_OCCTI_ANTS_14_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year
)[all_sp_ants]

occti_ants_all_sp_outputs <- filter_occupancy(occti_ants_all_sp_outputs, min_years_post_filter = 3)
```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_ants_all_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)
```

## Generate SPARTA posterior samples

NB: I am only the species that passed the occti filter

```{r}

sparta_ants_all_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(names(occti_ants_all_sp_outputs), collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year,
)[[1]]

```

## Compute BMA for all ant species

```{r}
# 1) bma plot sparta vs occti ants - they're different
# bma_sparta_ants_all_sp = calculate_indicator(sparta_ants_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_all_sp.rds")
# bma_occti_ants_all_sp = calculate_indicator(loess_results_ants_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_all_sp.rds")

bma_sparta_ants_all_sp = readRDS("bma_runs/bma_sparta_ants_all_sp.rds")
bma_occti_ants_all_sp = readRDS("bma_runs/bma_occti_ants_all_sp.rds")

sparta_plots_ants_all_sp <- summariseMSI(bma_sparta_ants_all_sp, label = "sparta all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_ants_all_sp <- summariseMSI(bma_occti_ants_all_sp,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_ants_all_sp$indicator + occti_plots_ants_all_sp$indicator
ggsave("bma_plots/ants_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

## Most common ant species

reset by removing CI and Z-score filters

```{r}
occti_ants_all_sp_outputs <- load_occti_outputs(
  results_dir = "past_jasmin_datalabs_runs/_rslurm_explore_occ_run_OCCTI_ANTS_14_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year
)

ants_most_common_sp = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
  rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
  count(species) %>%
  arrange(desc(n)) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_ants_most_common_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs[ants_most_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_ants_most_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(ants_most_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

# bma_sparta_ants_most_common_sp = calculate_indicator(sparta_ants_most_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_most_common_sp.rds")
# bma_occti_ants_most_common_sp = calculate_indicator(loess_results_ants_most_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_most_common_sp.rds")

bma_sparta_ants_most_common_sp = readRDS("bma_runs/bma_sparta_ants_most_common_sp.rds")
bma_occti_ants_most_common_sp = readRDS("bma_runs/bma_occti_ants_most_common_sp.rds")

sparta_plots_ants_most_common_sp <- summariseMSI(bma_sparta_ants_most_common_sp, label = "sparta most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_ants_most_common_sp <- summariseMSI(bma_occti_ants_most_common_sp,
                                      label = "occti most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_ants_most_common_sp$indicator + occti_plots_ants_most_common_sp$indicator
ggsave("bma_plots/ants_most_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

## Least common ant species

```{r}
ants_least_common_sp = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
count(species) %>%
arrange(n) %>%
filter(n > 50) %>%
slice(1:15) %>%
pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_ants_least_common_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs[ants_least_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_ants_least_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(ants_least_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

# bma_sparta_ants_least_common_sp = calculate_indicator(sparta_ants_least_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_least_common_sp.rds")
# bma_occti_ants_least_common_sp = calculate_indicator(loess_results_ants_least_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_least_common_sp.rds")

bma_sparta_ants_least_common_sp = readRDS("bma_runs/bma_sparta_ants_least_common_sp.rds")
bma_occti_ants_least_common_sp = readRDS("bma_runs/bma_occti_ants_least_common_sp.rds")

sparta_plots_ants_least_common_sp <- summariseMSI(bma_sparta_ants_least_common_sp, label = "sparta least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_ants_least_common_sp <- summariseMSI(bma_occti_ants_least_common_sp,
                                      label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_ants_least_common_sp$indicator + occti_plots_ants_least_common_sp$indicator
ggsave("bma_plots/ants_least_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

# Dragonflies - sparta and occti

## Load OCCTI outputs

```{r}
# Load raw data for later comparisons of most and least common species
dg_flies_data = read.csv("occurence_datasets/odonata_cleaned_data_2025.csv") %>% mutate(species = gsub("/", "_", species), 
date = ymd(date), year = year(date))

occti_dg_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_explore_occ_run_OCCTI_DRAGONFLIES_18_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)

names(occti_dg_flies_all_sp_outputs) = gsub("/", "_", names(occti_dg_flies_all_sp_outputs))

# Apply filter
occti_dg_flies_all_sp_outputs <- filter_occupancy(occti_dg_flies_all_sp_outputs)

# Compare species that overlap
length(names(occti_dg_flies_all_sp_outputs))
length(list.files("past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly", pattern = "\\.rdata$"))
```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_dg_flies_all_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## Generate SPARTA posterior samples

```{r}
sparta_dg_flies_all_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly",                       
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  write = FALSE,
  speciesToKeep = paste(names(occti_dg_flies_all_sp_outputs), collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]
```

## Compute BMA for all dragonflies

```{r}
# bma_sparta_dg_flies_all_sp = calculate_indicator(sparta_dg_flies_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_all_sp.rds")
# bma_occti_dg_flies_all_sp = calculate_indicator(loess_results_dg_flies_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_all_sp.rds")

bma_sparta_dg_flies_all_sp = readRDS("bma_runs/bma_sparta_dg_flies_all_sp.rds")
bma_occti_dg_flies_all_sp = readRDS("bma_runs/bma_occti_dg_flies_all_sp.rds")

sparta_plots_dg_flies_all_sp <- summariseMSI(bma_sparta_dg_flies_all_sp, label = "sparta all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_dg_flies_all_sp <- summariseMSI(bma_occti_dg_flies_all_sp,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_dg_flies_all_sp$indicator + occti_plots_dg_flies_all_sp$indicator
ggsave("bma_plots/dragonflies_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)

```


## Most common dragonflies species

```{r}
occti_dg_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_explore_occ_run_OCCTI_DRAGONFLIES_18_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)

dg_flies_most_common_sp = dg_flies_data %>%
  count(species) %>%
  arrange(desc(n)) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here occti_dg_flies_all_sp_outputs
loess_results_dg_flies_most_common_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs[dg_flies_most_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_dg_flies_most_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(dg_flies_most_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

# bma_sparta_dg_flies_most_common_sp = calculate_indicator(sparta_dg_flies_most_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_most_common_sp.rds")
# bma_occti_dg_flies_most_common_sp = calculate_indicator(loess_results_dg_flies_most_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_most_common_sp.rds")

bma_sparta_dg_flies_most_common_sp = readRDS("bma_runs/bma_sparta_dg_flies_most_common_sp.rds")
bma_occti_dg_flies_most_common_sp = readRDS("bma_runs/bma_occti_dg_flies_most_common_sp.rds")

sparta_plots_dg_flies_most_common_sp <- summariseMSI(bma_sparta_dg_flies_most_common_sp, label = "sparta most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_dg_flies_most_common_sp <- summariseMSI(bma_occti_dg_flies_most_common_sp,
                                      label = "occti most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_dg_flies_most_common_sp$indicator + occti_plots_dg_flies_most_common_sp$indicator
ggsave("bma_plots/dragonflies_most_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

## 1.6 Least common dragonfly species

```{r}

dg_flies_least_common_sp = dg_flies_data %>%
  count(species) %>%
  filter(n >= 50) %>%
  arrange(n) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_dg_flies_least_common_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs[dg_flies_least_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_dg_flies_least_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(dg_flies_least_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

# bma_sparta_dg_flies_least_common_sp = calculate_indicator(sparta_dg_flies_least_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_least_common_sp.rds")
# bma_occti_dg_flies_least_common_sp = calculate_indicator(loess_results_dg_flies_least_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_least_common_sp.rds")

bma_sparta_dg_flies_least_common_sp = readRDS("bma_runs/bma_sparta_dg_flies_least_common_sp.rds")
bma_occti_dg_flies_least_common_sp = readRDS("bma_runs/bma_occti_dg_flies_least_common_sp.rds")

sparta_plots_dg_flies_least_common_sp <- summariseMSI(bma_sparta_dg_flies_least_common_sp, label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_dg_flies_least_common_sp <- summariseMSI(bma_occti_dg_flies_least_common_sp,
                                      label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_dg_flies_least_common_sp$indicator + occti_plots_dg_flies_least_common_sp$indicator
ggsave("bma_plots/dragonflies_least_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

# Hoverflies - sparta and occti

## Load OCCTI outputs

```{r}
# Load raw data for later comparisons of most and least common species
hv_flies_data = read.csv("occurence_datasets/hoverfly_2025_cleaned_data_for_occti.csv") %>% mutate(species = gsub("/", "_", species), 
date = ymd(date), year = year(date))

occti_hv_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_explore_occ_run_OCCTI_HOVERFLIES_11_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)

names(occti_hv_flies_all_sp_outputs) = gsub("/", "_", names(occti_hv_flies_all_sp_outputs))

# Apply filter
occti_hv_flies_all_sp_outputs <- filter_occupancy(occti_hv_flies_all_sp_outputs)

# Compare species that overlap
length(names(occti_hv_flies_all_sp_outputs))
length(list.files("past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_Hoverfly", pattern = "\\.rdata$"))
```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_hv_flies_all_sp <- posterior_samp_occti(
  occti_outputs = occti_hv_flies_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## Generate SPARTA posterior samples

```{r}
sparta_hv_flies_all_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_hoverfly",                       
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  write = FALSE,
  speciesToKeep = paste(names(occti_hv_flies_all_sp_outputs), collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]
```

## Compute BMA for all hoverflies

```{r}
# bma_sparta_hv_flies_all_sp = calculate_indicator(sparta_hv_flies_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_hv_flies_all_sp.rds")
# bma_occti_hv_flies_all_sp = calculate_indicator(loess_results_hv_flies_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_hv_flies_all_sp.rds")

bma_sparta_hv_flies_all_sp = readRDS("bma_runs/bma_sparta_hv_flies_all_sp.rds")
bma_occti_hv_flies_all_sp = readRDS("bma_runs/bma_occti_hv_flies_all_sp.rds")

sparta_plots_hv_flies_all_sp <- summariseMSI(bma_sparta_hv_flies_all_sp, label = "sparta all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_hv_flies_all_sp <- summariseMSI(bma_occti_hv_flies_all_sp,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_hv_flies_all_sp$indicator + occti_plots_hv_flies_all_sp$indicator
ggsave("bma_plots/hoverflies_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)

```


## Most common hoverflies species

```{r}
occti_hv_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_explore_occ_run_OCCTI_HOVERFLIES_11_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)

hv_flies_most_common_sp = hv_flies_data %>%
  count(species) %>%
  arrange(desc(n)) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here occti_hv_flies_all_sp_outputs
loess_results_hv_flies_most_common_sp <- posterior_samp_occti(
  occti_outputs = occti_hv_flies_all_sp_outputs[hv_flies_most_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_hv_flies_most_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_hoverfly",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(hv_flies_most_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

# bma_sparta_hv_flies_most_common_sp = calculate_indicator(sparta_hv_flies_most_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_hv_flies_most_common_sp.rds")
# bma_occti_hv_flies_most_common_sp = calculate_indicator(loess_results_hv_flies_most_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_hv_flies_most_common_sp.rds")

bma_sparta_hv_flies_most_common_sp = readRDS("bma_runs/bma_sparta_hv_flies_most_common_sp.rds")
bma_occti_hv_flies_most_common_sp = readRDS("bma_runs/bma_occti_hv_flies_most_common_sp.rds")

sparta_plots_hv_flies_most_common_sp <- summariseMSI(bma_sparta_hv_flies_most_common_sp, label = "sparta most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_hv_flies_most_common_sp <- summariseMSI(bma_occti_hv_flies_most_common_sp,
                                      label = "occti most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_hv_flies_most_common_sp$indicator + occti_plots_hv_flies_most_common_sp$indicator
ggsave("bma_plots/hoverflies_most_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

## 1.6 Least common hoverflies species

```{r}

hv_flies_least_common_sp = hv_flies_data %>%
  count(species) %>%
  filter(n >= 50) %>%
  arrange(n) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_hv_flies_least_common_sp <- posterior_samp_occti(
  occti_outputs = occti_hv_flies_all_sp_outputs[hv_flies_least_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_hv_flies_least_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_hoverfly",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(hv_flies_least_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

# bma_sparta_hv_flies_least_common_sp = calculate_indicator(sparta_hv_flies_least_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_hv_flies_least_common_sp.rds")
# bma_occti_hv_flies_least_common_sp = calculate_indicator(loess_results_hv_flies_least_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_hv_flies_least_common_sp.rds")

bma_sparta_hv_flies_least_common_sp = readRDS("bma_runs/bma_sparta_hv_flies_least_common_sp.rds")
bma_occti_hv_flies_least_common_sp = readRDS("bma_runs/bma_occti_hv_flies_least_common_sp.rds")

sparta_plots_hv_flies_least_common_sp <- summariseMSI(bma_sparta_hv_flies_least_common_sp, label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_hv_flies_least_common_sp <- summariseMSI(bma_occti_hv_flies_least_common_sp,
                                      label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_hv_flies_least_common_sp$indicator + occti_plots_hv_flies_least_common_sp$indicator
ggsave("bma_plots/hoverflies_least_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

# Produce comparative plots for occti-sparta

Here, I produce comparisons for individual species. In this case I will compare methods for dragonflies and hoverflies

```{r}

produce_comparison_plots(loess_results_dg_flies_all_sp, "past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly", output_folder = "dg_flies_sparta_occti_comparison", sparta_region = NULL, year_min_for_occti_comp = trend_start_year, year_max_for_occti_comp = trend_end_year)
produce_comparison_plots(loess_results_dg_flies_most_common_sp, "past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly", output_folder = "dg_flies_most_common_sparta_occti_comparison", sparta_region = NULL, year_min_for_occti_comp = trend_start_year, year_max_for_occti_comp = trend_end_year)
produce_comparison_plots(loess_results_dg_flies_least_common_sp, "past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly", output_folder = "dg_flies_least_common_sparta_occti_comparison", sparta_region = NULL, year_min_for_occti_comp = trend_start_year, year_max_for_occti_comp = trend_end_year)

produce_comparison_plots(loess_results_hv_flies_all_sp, "past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_hoverfly", output_folder = "hv_flies_sparta_occti_comparison", sparta_region = NULL, year_min_for_occti_comp = trend_start_year, year_max_for_occti_comp = trend_end_year)
produce_comparison_plots(loess_results_hv_flies_most_common_sp, "past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_hoverfly", output_folder = "hv_flies_most_common_sparta_occti_comparison", sparta_region = NULL, year_min_for_occti_comp = trend_start_year, year_max_for_occti_comp = trend_end_year)
produce_comparison_plots(loess_results_hv_flies_least_common_sp, "past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_hoverfly", output_folder = "hv_flies_least_common_sparta_occti_comparison", sparta_region = NULL, year_min_for_occti_comp = trend_start_year, year_max_for_occti_comp = trend_end_year)

# produce_comparison_plots(loess_results_ants_all_sp, "past_jasmin_datalabs_runs/_rslurm_dylcar_explore_occ_run_SPARTA_ANTS_22_07_2025", output_folder = "ants_sparta_occti_comparison", sparta_region = NULL)
# produce_comparison_plots(loess_results_ants_most_common_sp, "past_jasmin_datalabs_runs/_rslurm_dylcar_explore_occ_run_SPARTA_ANTS_22_07_2025", output_folder = "ants_most_common_sparta_occti_comparison", sparta_region = NULL)
# produce_comparison_plots(loess_results_ants_least_common_sp, "past_jasmin_datalabs_runs/_rslurm_dylcar_explore_occ_run_SPARTA_ANTS_22_07_2025", output_folder = "ants_least_common_sparta_occti_comparison", sparta_region = NULL)

```

# Moths - occti

## Load OCCTI outputs

```{r}
# Load occti outputs
occti_moths_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/moths_sp/_rslurm_explore_occ_run_OCCTI_MOTHS_22_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)

names(occti_moths_all_sp_outputs) = gsub("/", "_", names(occti_moths_all_sp_outputs))

# Apply filter
occti_moths_all_sp_outputs <- filter_occupancy(occti_moths_all_sp_outputs)
```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_moths_all_sp <- posterior_samp_occti(
  occti_outputs = occti_moths_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## Compute BMA for all moths

```{r}
# bma_occti_moths_all_sp = calculate_indicator(loess_results_moths_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_moths_all_sp.rds")

bma_occti_moths_all_sp = readRDS("bma_runs/bma_occti_moths_all_sp.rds")

occti_plots_moths_all_sp <- summariseMSI(bma_occti_moths_all_sp,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = occti_plots_moths_all_sp$indicator
ggsave("bma_plots/moths_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)

```

# butterflies - occti

## Load OCCTI outputs

```{r}
# Load occti outputs
occti_butterflies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/Butterflies_sp/_rslurm_explore_occ_run_OCCTI_BUTTERFLIES_22_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)

names(occti_butterflies_all_sp_outputs) = gsub("/", "_", names(occti_butterflies_all_sp_outputs))

# Apply filter
occti_butterflies_all_sp_outputs <- filter_occupancy(occti_butterflies_all_sp_outputs)
```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_butterflies_all_sp <- posterior_samp_occti(
  occti_outputs = occti_butterflies_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## Compute BMA for all butterflies

```{r}
# bma_occti_butterflies_all_sp = calculate_indicator(loess_results_butterflies_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_butterflies_all_sp.rds")

bma_occti_butterflies_all_sp = readRDS("bma_runs/bma_occti_butterflies_all_sp.rds")

occti_plots_butterflies_all_sp <- summariseMSI(bma_occti_butterflies_all_sp,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = occti_plots_butterflies_all_sp$indicator
ggsave("bma_plots/butterflies_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)

```

# Generate some new occupancy trends for comparison
```{r}
library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)
library(sparta)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

sparta_dirs = c("past_jasmin_datalabs_runs/Craneflies")
taxon_names = c("craneflies")
# sparta_dirs = c("past_jasmin_datalabs_runs/_rslurm_hoverfly", "past_jasmin_datalabs_runs/_rslurm_Dragonfly")
# taxon_names = c("hoverflies", "dragonflies")

# Trend start year
trend_start_year = 1980
trend_end_year = 2024

plots = list()

for (i in seq_along(sparta_dirs)){

  dir_i = sparta_dirs[i]
  files = list.files(dir_i, pattern = ".rdata", full.names = TRUE)
  taxon_group = taxon_names[i]

  for (file in files){
    species = gsub(".rdata", "", basename(file))
    species = gsub("/", "_", species)

    load(file)

    ggsave(file.path("comparative_occ_plots", taxon_group, paste0(species, ".png")), plot = plot(out), width = 8, height = 6, dpi = 300)

  }

}

```

# Combining bryophyte daisy chains

```{r}
library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

combine_daisy_chains("test_daisy_chains", output_path = "test_combined_chains")

trend_start_year = 1980

# Test other functionality
sparta_bryophytes_all_sp_posterior_samples <- applySamp_single(
  modPath = "test_combined_chains",
  region = "ENGLAND",
  nSamps = 999,
  minObs = NULL,
  t0 = 1980,
  tn = 2025,
)

filter_rot(sparta_bryophytes_all_sp_posterior_samples, region = "ENGLAND")

# bma_sparta_bryophytes_all_sp = calculate_indicator(sparta_bryophytes_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_bryophytes_all_sp.rds")

```

```{r}
library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

#test short term indicator
# calculate_indicator(sparta_ants_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", calculate_st_trends = TRUE, outputs_path = "bma_runs/bma_sparta_ants_all_sp_str_trends.rds")
```

# Combining posterior samples for England Indicator

```{r}
library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

occ_dirs = c("past_jasmin_datalabs_runs/Hoverflies_Sp/_rslurm_hoverfly",
"past_jasmin_datalabs_runs/Butterflies_sp/_rslurm_explore_occ_run_OCCTI_BUTTERFLIES_22_08_2025/results",
"past_jasmin_datalabs_runs/moths_sp/_rslurm_explore_occ_run_OCCTI_MOTHS_22_08_2025/results",
"past_jasmin_datalabs_runs/Odonata_Sp/_rslurm_Dragonfly")
sparta_used = c(TRUE, FALSE, FALSE, TRUE)

# Trend start year
trend_start_year = 1980
trend_end_year = 2024

all_posterior_samps = list()

for (i in seq_along(occ_dirs)){

  dir_i = occ_dirs[i]
  sparta_i = sparta_used[i]

  print(dir_i)
  print(sparta_i)

  if(sparta_i){

  # posterior sample sparta run outputs
  sparta_posterior_samps <- applySamp_single(
    modPath = dir_i,
    region = "UK",
    nSamps = 999,
    minObs = NULL,
    speciesToKeep = NA,
    t0 = trend_start_year,
    tn = trend_end_year,
  )

  sparta_posterior_samps <- filter_rot(sparta_posterior_samps)

  # Translate the wide posterior samples to a long format
  sparta_posterior_samps = translate_sparta(sparta_posterior_samps$posterior_samples)

  # NB: we need to do species list filtering here

  all_posterior_samps[[i]] = sparta_posterior_samps

  } else{

  occti_outputs <- load_occti_outputs(
    results_dir = dir_i,
    pattern = "uk_occupancy_output.rds",
    start_year = trend_start_year,
    end_year = trend_end_year
  )

  # Apply filter
  occti_outputs <- filter_occupancy(occti_outputs)

  # Obtain posterior occti samples
  posterior_samps <- posterior_samp_occti(
    occti_outputs = occti_outputs,
    save_plots = FALSE,
    span = 0.75,
    plot_width = 18,
    plot_height = 6,
    n_iter = 1000
  )

  # NB: we need to do species list filtering here

  all_posterior_samps[[i]] = posterior_samps

  }

}

saveRDS(all_posterior_samps, "all_groups_posterior_samps.RDS")

# all_bma = calculate_indicator(all_posterior_samps, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = FALSE)

all_bma_plots <- summariseMSI(all_bma,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

```

```{r}

library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)
library(lubridate)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

# paper stats
# most common dg_fly species
data = read.csv("occurence_datasets/odonata_cleaned_data_2025.csv")
data %>% group_by(species) %>% summarise(n_records = n()) %>% filter(n_records >= 50) %>% arrange(n_records)

# most common hv_fly species
read.csv("occurence_datasets/hoverfly_2025_cleaned_data_for_occti.csv") %>% mutate(species = gsub("/", "_", species), date = ymd(date), year = year(date)) %>% group_by(species) %>%
summarise(n_records = n()) %>% filter(n_records >= 50) %>% arrange(n_records)

# Row count
nrow(read.csv("occurence_datasets/BNM_data_1970to2020_180825.csv"))
nrow(read.csv("occurence_datasets/NMRS_data_1970to2020_180825.csv"))

```