---
title: "Multi‑Species Indicator (MSI) Workflow"
author: "Dylan"
output: HTML
---

# Setup

```{r}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)

library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

# Trend start year
trend_start_year = 1970
trend_end_year = 2024

```

# Ants - sparta and occti

## Load OCCTI outputs

```{r}

# Forgot to implement 50 record filter.
all_sp_ants = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
count(species) %>%
filter(n >= 50) %>%
pull(species)

occti_ants_all_sp_outputs <- load_occti_outputs(
  results_dir = "past_jasmin_datalabs_runs/_rslurm_explore_occ_run_OCCTI_ANTS_14_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year
)[all_sp_ants]

occti_ants_all_sp_outputs <- filter_occupancy(occti_ants_all_sp_outputs, min_years_post_filter = 3)
```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_ants_all_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)
```

## Generate SPARTA posterior samples

NB: I am only the species that passed the occti filter

```{r}

sparta_ants_all_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(names(occti_ants_all_sp_outputs), collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year,
)[[1]]

```

## Compute BMA for all ant species

```{r}
# 1) bma plot sparta vs occti ants - they're different
bma_sparta_ants_all_sp = calculate_indicator(sparta_ants_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_all_sp.rds")
bma_occti_ants_all_sp = calculate_indicator(loess_results_ants_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_all_sp.rds")

bma_sparta_ants_all_sp = readRDS("bma_runs/bma_sparta_ants_all_sp.rds")
bma_occti_ants_all_sp = readRDS("bma_runs/bma_occti_ants_all_sp.rds")

sparta_plots_ants_all_sp <- summariseMSI(bma_sparta_ants_all_sp, label = "sparta all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_ants_all_sp <- summariseMSI(bma_occti_ants_all_sp,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_ants_all_sp$indicator + occti_plots_ants_all_sp$indicator
ggsave("bma_plots/ants_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

## Most common ant species

reset by removing CI and Z-score filters

```{r}
occti_ants_all_sp_outputs <- load_occti_outputs(
  results_dir = "past_jasmin_datalabs_runs/_rslurm_explore_occ_run_OCCTI_ANTS_14_08_2025/results",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year
)

ants_most_common_sp = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
  rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
  count(species) %>%
  arrange(desc(n)) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_ants_most_common_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs[ants_most_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_ants_most_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(ants_most_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

bma_sparta_ants_most_common_sp = calculate_indicator(sparta_ants_most_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_most_common_sp.rds")
bma_occti_ants_most_common_sp = calculate_indicator(loess_results_ants_most_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_most_common_sp.rds")

bma_sparta_ants_most_common_sp = readRDS("bma_runs/bma_sparta_ants_most_common_sp.rds")
bma_occti_ants_most_common_sp = readRDS("bma_runs/bma_occti_ants_most_common_sp.rds")

sparta_plots_ants_most_common_sp <- summariseMSI(bma_sparta_ants_most_common_sp, label = "sparta most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_ants_most_common_sp <- summariseMSI(bma_occti_ants_most_common_sp,
                                      label = "occti most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_ants_most_common_sp$indicator + occti_plots_ants_most_common_sp$indicator
ggsave("bma_plots/ants_most_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

## Least common ant species

```{r}
ants_least_common_sp = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
count(species) %>%
arrange(n) %>%
filter(n > 50) %>%
slice(1:15) %>%
pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_ants_least_common_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs[ants_least_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_ants_least_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(ants_least_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

bma_sparta_ants_least_common_sp = calculate_indicator(sparta_ants_least_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_least_common_sp.rds")
bma_occti_ants_least_common_sp = calculate_indicator(loess_results_ants_least_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_least_common_sp.rds")

bma_sparta_ants_least_common_sp = readRDS("bma_runs/bma_sparta_ants_least_common_sp.rds")
bma_occti_ants_least_common_sp = readRDS("bma_runs/bma_occti_ants_least_common_sp.rds")

sparta_plots_ants_least_common_sp <- summariseMSI(bma_sparta_ants_least_common_sp, label = "sparta least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_ants_least_common_sp <- summariseMSI(bma_occti_ants_least_common_sp,
                                      label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_ants_least_common_sp$indicator + occti_plots_ants_least_common_sp$indicator
ggsave("bma_plots/ants_least_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

# Dragonflies - sparta and occti

## Load OCCTI outputs

```{r}
# Load raw data for later comparisons of most and least common species
load("occurence_datasets/odonata_2022.rdata")

dg_flies_data = visitData$spp_vis %>%
  pivot_longer(
    cols = -visit,
    names_to = "species",
    values_to = "present"
  ) %>%
  filter(present) %>%
  select(-present) %>%
  mutate(
    gridref = str_extract(visit, "^[A-Z]{2}\\d{4}"),
    date = str_extract(visit, "\\d{4}-\\d{2}-\\d{2}") %>% as.Date()
  ) %>%
  select(-visit)

all_sp_dg_flies = dg_flies_data %>%
count(species) %>%
filter(n >= 50) %>%
pull(species)

occti_dg_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/dragonflies_results_12_08_2025",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)[all_sp_dg_flies]

# Apply filter
occti_dg_flies_all_sp_outputs <- filter_occupancy(occti_dg_flies_all_sp_outputs)

# Compare species that overlap
length(names(occti_dg_flies_all_sp_outputs))
length(list.files("past_runs_sparta/Odonata_2022_Powney"))
```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_dg_flies_all_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## Generate SPARTA posterior samples

```{r}
sparta_dg_flies_all_sp_posterior_samples <- applySamp_single(
  modPath = "past_runs_sparta/Odonata_2022_Powney",                       
  region = NULL,
  nSamps = 999,
  minObs = NULL,
  write = FALSE,
  speciesToKeep = paste(names(occti_dg_flies_all_sp_outputs), collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]
```

## Compute BMA for all dragonflies

```{r}
bma_sparta_dg_flies_all_sp = calculate_indicator(sparta_dg_flies_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_all_sp.rds")
bma_occti_dg_flies_all_sp = calculate_indicator(loess_results_dg_flies_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_all_sp.rds")

bma_sparta_dg_flies_all_sp = readRDS("bma_runs/bma_sparta_dg_flies_all_sp.rds")
bma_occti_dg_flies_all_sp = readRDS("bma_runs/bma_occti_dg_flies_all_sp.rds")

sparta_plots_dg_flies_all_sp <- summariseMSI(bma_sparta_dg_flies_all_sp, label = "sparta all species",
                                      minYear = 1970,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_dg_flies_all_sp <- summariseMSI(bma_occti_dg_flies_all_sp,
                                      label = "occti all species",
                                      minYear = 1970,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_dg_flies_all_sp$indicator + occti_plots_dg_flies_all_sp$indicator
ggsave("bma_plots/dragonflies_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)

```


## Most common dragonflies species

```{r}
occti_dg_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/dragonflies_results_12_08_2025",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)

dg_flies_most_common_sp = dg_flies_data %>%
  count(species) %>%
  arrange(desc(n)) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here occti_dg_flies_all_sp_outputs
loess_results_dg_flies_most_common_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs[dg_flies_most_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_dg_flies_most_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_runs_sparta/Odonata_2022_Powney",        
  region = NULL,
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(dg_flies_most_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

bma_sparta_dg_flies_most_common_sp = calculate_indicator(sparta_dg_flies_most_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_most_common_sp.rds")
bma_occti_dg_flies_most_common_sp = calculate_indicator(loess_results_dg_flies_most_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_most_common_sp.rds")

bma_sparta_dg_flies_most_common_sp = readRDS("bma_runs/bma_sparta_dg_flies_most_common_sp.rds")
bma_occti_dg_flies_most_common_sp = readRDS("bma_runs/bma_occti_dg_flies_most_common_sp.rds")

sparta_plots_dg_flies_most_common_sp <- summariseMSI(bma_sparta_dg_flies_most_common_sp, label = "sparta most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_dg_flies_most_common_sp <- summariseMSI(bma_occti_dg_flies_most_common_sp,
                                      label = "occti most common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_dg_flies_most_common_sp$indicator + occti_plots_dg_flies_most_common_sp$indicator
ggsave("bma_plots/dragonflies_most_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

## 1.6 Least common dragonfly species

```{r}

dg_flies_least_common_sp = dg_flies_data %>%
  count(species) %>%
  filter(n >= 50) %>%
  arrange(n) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_dg_flies_least_common_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs[dg_flies_least_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_dg_flies_least_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_runs_sparta/Odonata_2022_Powney",        
  region = NULL,
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(dg_flies_least_common_sp, collapse = ","),
  t0 = trend_start_year,
  tn = trend_end_year)[[1]]

bma_sparta_dg_flies_least_common_sp = calculate_indicator(sparta_dg_flies_least_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_least_common_sp.rds")
bma_occti_dg_flies_least_common_sp = calculate_indicator(loess_results_dg_flies_least_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_least_common_sp.rds")

bma_sparta_dg_flies_least_common_sp = readRDS("bma_runs/bma_sparta_dg_flies_least_common_sp.rds")
bma_occti_dg_flies_least_common_sp = readRDS("bma_runs/bma_occti_dg_flies_least_common_sp.rds")

sparta_plots_dg_flies_least_common_sp <- summariseMSI(bma_sparta_dg_flies_least_common_sp, label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

occti_plots_dg_flies_least_common_sp <- summariseMSI(bma_occti_dg_flies_least_common_sp,
                                      label = "occti least common species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = sparta_plots_dg_flies_least_common_sp$indicator + occti_plots_dg_flies_least_common_sp$indicator
ggsave("bma_plots/dragonflies_least_common_sp.png", plot = plot, width = 16, height = 4, dpi = 300)
```

# Produce comparative plots for occti-sparta

Here, I produce comparisons for individual species. I start with dragonflies.
```{r}

produce_comparison_plots(loess_results_dg_flies_all_sp, "past_runs_sparta/Odonata_2022_Powney", output_folder = "dg_flies_sparta_occti_comparison", sparta_region = NULL)
produce_comparison_plots(loess_results_dg_flies_most_common_sp, "past_runs_sparta/Odonata_2022_Powney", output_folder = "dg_flies_most_common_sparta_occti_comparison", sparta_region = NULL)
produce_comparison_plots(loess_results_dg_flies_least_common_sp, "past_runs_sparta/Odonata_2022_Powney", output_folder = "dg_flies_least_common_sparta_occti_comparison", sparta_region = NULL)

produce_comparison_plots(loess_results_ants_all_sp, "past_jasmin_datalabs_runs/_rslurm_dylcar_explore_occ_run_SPARTA_ANTS_22_07_2025", output_folder = "ants_sparta_occti_comparison", sparta_region = NULL)
produce_comparison_plots(loess_results_ants_most_common_sp, "past_jasmin_datalabs_runs/_rslurm_dylcar_explore_occ_run_SPARTA_ANTS_22_07_2025", output_folder = "ants_most_common_sparta_occti_comparison", sparta_region = NULL)
produce_comparison_plots(loess_results_ants_least_common_sp, "past_jasmin_datalabs_runs/_rslurm_dylcar_explore_occ_run_SPARTA_ANTS_22_07_2025", output_folder = "ants_least_common_sparta_occti_comparison", sparta_region = NULL)

```

# Hoverflies - occti

## Load OCCTI outputs

```{r}

all_sp_hv_flies = read.csv("occurence_datasets/hoverfly_2025_cleaned_data_for_occti.csv") %>%
count(species) %>%
filter(n >= 50) %>%
pull(species)

occti_hv_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/hoverflies_results_11_08_2025",
  pattern = "uk_occupancy_output.rds",
  start_year = trend_start_year,
  end_year = trend_end_year)[all_sp_hv_flies]

occti_hv_flies_all_sp_outputs = occti_hv_flies_all_sp_outputs[!is.na(names(occti_hv_flies_all_sp_outputs))] # "Melangyna compositarum/labiatarum" failed to run. Removing as NA

# Apply filter
occti_hv_flies_all_sp_outputs <- filter_occupancy(occti_hv_flies_all_sp_outputs)

```

## Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_hv_flies_all_sp <- posterior_samp_occti(
  occti_outputs = occti_hv_flies_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## Compute BMA for all hoverflies

```{r}
bma_occti_hv_flies_all_sp = calculate_indicator(loess_results_hv_flies_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_hv_flies_all_sp.rds")

bma_occti_hv_flies_all_sp = readRDS("bma_runs/bma_occti_hv_flies_all_sp.rds")

occti_plots_hv_flies_all_sp <- summariseMSI(bma_occti_hv_flies_all_sp,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plot = occti_plots_hv_flies_all_sp$indicator
ggsave("bma_plots/hoverflies_all_sp.png", plot = plot, width = 16, height = 4, dpi = 300)

```

# Rovebeetles - sparta

```{r}

library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

# Trend start year
trend_start_year = 1970
trend_end_year = 2024

# I am going to load the posterior samples - They prirably need to be stored locally as accessing them from the shared folder can take very long
# NB: The species output are all lowercase and this may affect merging. Please let me know if I should change this
sparta_rv_beetles_all_sp_posterior_samples <- applySamp_single(
  modPath = "RoveBeetles_Sp/_rslurm_RoveBeetles",
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = NA,
  t0 = trend_start_year,
  tn = trend_end_year,
)

# Somewhere here we need to filter based on a species list...

sparta_rv_beetles_all_sp_posterior_samples <- filter_rot(sparta_rv_beetles_all_sp_posterior_samples)

# Produce an indicator
bma_sparta_rv_beetles_all_sp = calculate_indicator(sparta_rv_beetles_all_sp_posterior_samples$posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_rv_beetles_all_sp.rds")

```

# Generate some new occupancy trends for comparison
```{r}
library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

sparta_dirs = c("past_jasmin_datalabs_runs/_rslurm_hoverfly", "past_jasmin_datalabs_runs/_rslurm_Dragonfly")
taxon_names = c("hoverflies", "dragonflies")

# Trend start year
trend_start_year = 1970
trend_end_year = 2024

plots = list()

for (i in seq_along(sparta_dirs)){

  dir_i = sparta_dirs[i]
  taxon_group = taxon_names[i]

  # posterior sample sparta run outputs
  sparta_posterior_samps <- applySamp_single(
    modPath = dir_i,
    region = "UK",
    nSamps = 999,
    minObs = NULL,
    speciesToKeep = NA,
    t0 = trend_start_year,
    tn = trend_end_year,
  )

sparta_posterior_samps <- filter_rot(sparta_posterior_samps)

bma_output = calculate_indicator(sparta_posterior_samps$posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = FALSE)

plot <- summariseMSI(bma_output,
                                      label = paste0("sparta_taxon_group_", taxon_group),
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

plots[[taxon_group]] = plot

}

saveRDS(plots, "bma_runs_dg_flies_hv_flies.rds")

```


# Combining posterior samples for England Indicator

```{r}
library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

occ_dirs = c()
sparta_used = c(TRUE, FALSE)

# Trend start year
trend_start_year = 1970
trend_end_year = 2024

all_posterior_samps = list()

for (i in seq_along(occ_dirs)){

  dir_i = occ_dirs[i]
  sparta_i = sparta_used[i]

  print(dir_i)
  print(sparta_i)

  if(sparta_i){

  # posterior sample sparta run outputs
  sparta_posterior_samps <- applySamp_single(
    modPath = dir_i,
    region = "UK",
    nSamps = 999,
    minObs = NULL,
    speciesToKeep = NA,
    t0 = trend_start_year,
    tn = trend_end_year,
  )

  sparta_posterior_samps <- filter_rot(sparta_posterior_samps)

  # Translate the wide posterior samples to a long format
  sparta_posterior_samps = translate_sparta(sparta_posterior_samps)

  all_posterior_samps[[i]] = sparta_posterior_samps

  } else{

  occti_outputs <- load_occti_outputs(
    results_dir = dir_i,
    pattern = "uk_occupancy_output.rds",
    start_year = trend_start_year,
    end_year = trend_end_year
  )

  # Apply filter
  occti_outputs <- filter_occupancy(occti_outputs)

  # Obtain posterior occti samples
  posterior_samps <- posterior_samp_occti(
    occti_outputs = occti_outputs,
    save_plots = FALSE,
    span = 0.75,
    plot_width = 18,
    plot_height = 6,
    n_iter = 1000
  )

  # NB: we need to do species list filtering here

  all_posterior_samps[[i]] = sparta_posterior_samp

  }

}

saveRDS(all_posterior_samps, "all_groups_posterior_samps.RDS")

all_bma = calculate_indicator(all_posterior_samps, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = FALSE)

all_bma_plots <- summariseMSI(all_bma,
                                      label = "occti all species",
                                      minYear = trend_start_year,
                                      maxYear = trend_end_year,
                                      writePlot = TRUE)

```
