---
title: "Multi‑Species Indicator (MSI) Workflow"
author: "Dylan"
output: HTML
---

# 0) Setup

```{r}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)

library(dplyr)
library(devtools)
library(BRCindicators)
library(tidyr)
library(patchwork)
library(stringr)
library(tidyr)

# use local package helpers (e.g., posterior_samp_occti, load_occti_outputs)
load_all("../occLite")

```

# 1) Ants (Formicidae) comparison with sparta

## 1.1 Load OCCTI outputs

```{r}

# Forgot to implement 50 record filter.
all_sp_ants = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
count(species) %>%
filter(n >= 50) %>%
pull(species)

occti_ants_all_sp_outputs <- load_occti_outputs(
  results_dir = "past_jasmin_datalabs_runs/_rslurm_explore_occ_run_OCCTI_ANTS_14_08_2025/results",
  pattern = "uk_occupancy_output.rds"
)[all_sp_ants]

occti_ants_all_sp_outputs <- filter_occupancy(occti_ants_all_sp_outputs, min_years_post_filter = 3)
```

## 1.2 Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_ants_all_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)
```

## 1.3 Generate SPARTA posterior samples

NB: I am only the species that passed the occti filter

```{r}
sparta_ants_all_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(names(occti_ants_all_sp_outputs), collapse = ","),
  t0 = 2000,
  tn = 2020
)[[1]]
```

## 1.4 Compute MSI (BMA) for all ant species

```{r}
# 1) bma plot sparta vs occti ants - they're different
bma_sparta_ants_all_sp = calculate_indicator(sparta_ants_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_all_sp.rds")
bma_occti_ants_all_sp = calculate_indicator(loess_results_ants_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_all_sp.rds")

bma_sparta_ants_all_sp = readRDS("bma_runs/bma_sparta_ants_all_sp.rds")
bma_occti_ants_all_sp = readRDS("bma_runs/bma_occti_ants_all_sp.rds")

sparta_plots_ants_all_sp <- summariseMSI(bma_sparta_ants_all_sp, label = "sparta all species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

occti_plots_ants_all_sp <- summariseMSI(bma_occti_ants_all_sp,
                                      label = "occti all species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

sparta_plots_ants_all_sp$indicator + occti_plots_ants_all_sp$indicator
```


## reset by removing CI and Z-score filters
```{r}
occti_ants_all_sp_outputs <- load_occti_outputs(
  results_dir = "past_jasmin_datalabs_runs/_rslurm_explore_occ_run_OCCTI_ANTS_14_08_2025/results"
)

```

## 1.5 Most common ant species

```{r}
ants_most_common_sp = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
  rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
  count(species) %>%
  arrange(desc(n)) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_ants_most_common_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs[ants_most_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_ants_most_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(ants_most_common_sp, collapse = ","),
  t0 = 2000,
  tn = 2020)[[1]]

bma_sparta_ants_most_common_sp = calculate_indicator(sparta_ants_most_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_most_common_sp.rds")
bma_occti_ants_most_common_sp = calculate_indicator(loess_results_ants_most_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_most_common_sp.rds")

bma_sparta_ants_most_common_sp = readRDS("bma_runs/bma_sparta_ants_most_common_sp.rds")
bma_occti_ants_most_common_sp = readRDS("bma_runs/bma_occti_ants_most_common_sp.rds")

sparta_plots_ants_most_common_sp <- summariseMSI(bma_sparta_ants_most_common_sp, label = "sparta most common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

occti_plots_ants_most_common_sp <- summariseMSI(bma_occti_ants_most_common_sp,
                                      label = "occti most common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

sparta_plots_ants_most_common_sp$indicator + occti_plots_ants_most_common_sp$indicator
```

## 1.6 Least common ant species

```{r}
ants_least_common_sp = readRDS("occurence_datasets/monad_occupancy_dataset_ants.rds") %>%
rename(date = lower_date, gridref = GRIDREF, species = tik) %>%
count(species) %>%
arrange(n) %>%
filter(n > 50) %>%
slice(1:15) %>%
pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_ants_least_common_sp <- posterior_samp_occti(
  occti_outputs = occti_ants_all_sp_outputs[ants_least_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_ants_least_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_jasmin_datalabs_runs/datalabs_ants_sparta_RW",        
  region = "UK",
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(ants_least_common_sp, collapse = ","),
  t0 = 2000,
  tn = 2020)[[1]]

bma_sparta_ants_least_common_sp = calculate_indicator(sparta_ants_least_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_ants_least_common_sp.rds")
bma_occti_ants_least_common_sp = calculate_indicator(loess_results_ants_least_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_ants_least_common_sp.rds")

bma_sparta_ants_least_common_sp = readRDS("bma_runs/bma_sparta_ants_least_common_sp.rds")
bma_occti_ants_least_common_sp = readRDS("bma_runs/bma_occti_ants_least_common_sp.rds")

sparta_plots_ants_least_common_sp <- summariseMSI(bma_sparta_ants_least_common_sp, label = "sparta least common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

occti_plots_ants_least_common_sp <- summariseMSI(bma_occti_ants_least_common_sp,
                                      label = "occti least common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)
```

# 2) Dragonflies (Odonata)

## 2.1 Load OCCTI outputs

```{r}
# Load raw data for later comparisons of most and least common species
load("occurence_datasets/odonata_2022.rdata")

dg_flies_data = visitData$spp_vis %>%
  pivot_longer(
    cols = -visit,
    names_to = "species",
    values_to = "present"
  ) %>%
  filter(present) %>%
  select(-present) %>%
  mutate(
    gridref = str_extract(visit, "^[A-Z]{2}\\d{4}"),
    date = str_extract(visit, "\\d{4}-\\d{2}-\\d{2}") %>% as.Date()
  ) %>%
  select(-visit)

all_sp_dg_flies = dg_flies_data %>%
count(species) %>%
filter(n >= 50) %>%
pull(species)

occti_dg_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/dragonflies_results_12_08_2025",
pattern = "uk_occupancy_output.rds")[all_sp_dg_flies]

# Apply filter
occti_dg_flies_all_sp_outputs <- filter_occupancy(occti_dg_flies_all_sp_outputs)

# Compare species that overlap
length(names(occti_dg_flies_all_sp_outputs))
length(list.files("past_runs_sparta/Odonata_2022_Powney"))
```

## 2.2 Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_dg_flies_all_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## 2.3 Generate SPARTA posterior samples

```{r}
sparta_dg_flies_all_sp_posterior_samples <- applySamp_single(
  modPath = "past_runs_sparta/Odonata_2022_Powney",                       
  region = NULL,
  nSamps = 999,
  minObs = NULL,
  write = FALSE,
  speciesToKeep = paste(names(occti_dg_flies_all_sp_outputs), collapse = ","),
  t0 = 2000,
  tn = 2020)[[1]]
```

## 2.4 Compute MSI (BMA) for all dragonflies

```{r}
bma_sparta_dg_flies_all_sp = calculate_indicator(sparta_dg_flies_all_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_all_sp.rds")
bma_occti_dg_flies_all_sp = calculate_indicator(loess_results_dg_flies_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_all_sp.rds")

bma_sparta_dg_flies_all_sp = readRDS("bma_runs/bma_sparta_dg_flies_all_sp.rds")
bma_occti_dg_flies_all_sp = readRDS("bma_runs/bma_occti_dg_flies_all_sp.rds")

sparta_plots_dg_flies_all_sp <- summariseMSI(bma_sparta_dg_flies_all_sp, label = "sparta all species",
                                      minYear = 1970,
                                      maxYear = 2022,
                                      writePlot = TRUE)

occti_plots_dg_flies_all_sp <- summariseMSI(bma_occti_dg_flies_all_sp,
                                      label = "occti all species",
                                      minYear = 1970,
                                      maxYear = 2022,
                                      writePlot = TRUE)

sparta_plots_dg_flies_all_sp$indicator + occti_plots_dg_flies_all_sp$indicator

```


## 1.5 Most common dragonflies species

```{r}
dg_flies_most_common_sp = dg_flies_data %>%
  count(species) %>%
  arrange(desc(n)) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here occti_dg_flies_all_sp_outputs
loess_results_dg_flies_most_common_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs[dg_flies_most_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_dg_flies_most_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_runs_sparta/Odonata_2022_Powney",        
  region = NULL,
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(dg_flies_most_common_sp, collapse = ","),
  t0 = 2000,
  tn = 2020)[[1]]

bma_sparta_dg_flies_most_common_sp = calculate_indicator(sparta_dg_flies_most_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_most_common_sp.rds")
bma_occti_dg_flies_most_common_sp = calculate_indicator(loess_results_dg_flies_most_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_most_common_sp.rds")

bma_sparta_dg_flies_most_common_sp = readRDS("bma_runs/bma_sparta_dg_flies_most_common_sp.rds")
bma_occti_dg_flies_most_common_sp = readRDS("bma_runs/bma_occti_dg_flies_most_common_sp.rds")

sparta_plots_dg_flies_most_common_sp <- summariseMSI(bma_sparta_dg_flies_most_common_sp, label = "sparta most common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

occti_plots_dg_flies_most_common_sp <- summariseMSI(bma_occti_dg_flies_most_common_sp,
                                      label = "occti most common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)
```

## 1.6 Least common dragonfly species

```{r}

dg_flies_least_common_sp = dg_flies_data %>%
  count(species) %>%
  filter(n >= 50) %>%
  arrange(n) %>%
  slice(1:15) %>%
  pull(species)

# Obtain posterior occti samples
# NB: I have not filtered the data here
loess_results_dg_flies_least_common_sp <- posterior_samp_occti(
  occti_outputs = occti_dg_flies_all_sp_outputs[dg_flies_least_common_sp],
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

# posterior sample sparta run outputs
sparta_dg_flies_least_common_sp_posterior_samples <- applySamp_single(
  modPath = "past_runs_sparta/Odonata_2022_Powney",        
  region = NULL,
  nSamps = 999,
  minObs = NULL,
  speciesToKeep = paste(dg_flies_least_common_sp, collapse = ","),
  t0 = 2000,
  tn = 2020)[[1]]

bma_sparta_dg_flies_least_common_sp = calculate_indicator(sparta_dg_flies_least_common_sp_posterior_samples, sparta_output = TRUE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_sparta_dg_flies_least_common_sp.rds")
bma_occti_dg_flies_least_common_sp = calculate_indicator(loess_results_dg_flies_least_common_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_dg_flies_least_common_sp.rds")

bma_sparta_dg_flies_least_common_sp = readRDS("bma_runs/bma_sparta_dg_flies_least_common_sp.rds")
bma_occti_dg_flies_least_common_sp = readRDS("bma_runs/bma_occti_dg_flies_least_common_sp.rds")

sparta_plots_dg_flies_least_common_sp <- summariseMSI(bma_sparta_dg_flies_least_common_sp, label = "occti least common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

occti_plots_dg_flies_least_common_sp <- summariseMSI(bma_occti_dg_flies_least_common_sp,
                                      label = "occti least common species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)
```

# 2) Hoverflies

## 2.1 Load OCCTI outputs

```{r}

all_sp_hv_flies = read.csv("occurence_datasets/hoverfly_2025_cleaned_data_for_occti.csv") %>%
count(species) %>%
filter(n >= 50) %>%
pull(species)

occti_hv_flies_all_sp_outputs <- load_occti_outputs("past_jasmin_datalabs_runs/hoverflies_results_11_08_2025",
pattern = "uk_occupancy_output.rds")[all_sp_hv_flies] 
occti_hv_flies_all_sp_outputs = occti_hv_flies_all_sp_outputs[!is.na(names(occti_hv_flies_all_sp_outputs))] # "Melangyna compositarum/labiatarum" failed to run. Removing as NA

# Apply filter
occti_hv_flies_all_sp_outputs <- filter_occupancy(occti_hv_flies_all_sp_outputs)

```

## 2.2 Generate OCCTI posterior samples

```{r}
# Obtain posterior occti samples
loess_results_hv_flies_all_sp <- posterior_samp_occti(
  occti_outputs = occti_hv_flies_all_sp_outputs,
  save_plots = FALSE,
  span = 0.75,
  plot_width = 18,
  plot_height = 6,
  n_iter = 1000
)

```

## 2.3 Generate SPARTA posterior samples

## 2.4 Compute MSI (BMA) for all hoverflies

```{r}
bma_occti_hv_flies_all_sp = calculate_indicator(loess_results_hv_flies_all_sp, sparta_output = FALSE, method = "bma", bma_ind = "prime", save_outputs = TRUE, outputs_path = "bma_runs/bma_occti_hv_flies_all_sp.rds")

bma_occti_hv_flies_all_sp = readRDS("bma_runs/bma_occti_hv_flies_all_sp.rds")

occti_plots_hv_flies_all_sp <- summariseMSI(bma_occti_hv_flies_all_sp,
                                      label = "occti all species",
                                      minYear = 1980,
                                      maxYear = 2022,
                                      writePlot = TRUE)

```

# 3) comparisons

Here, I produce comparisons for individual species. I start with dragonflies.
```{r}

load_all("../occLite")

produce_comparison_plots(loess_results_dg_flies_all_sp, "past_runs_sparta/Odonata_2022_Powney", output_folder = "dg_flies_sparta_occti_comparison", sparta_region = NULL)
produce_comparison_plots(loess_results_dg_flies_all_sp, "past_runs_sparta/Odonata_2022_Powney", output_folder = "dg_flies_sparta_occti_comparison", sparta_region = "UK")


```